Streaming $2$-Center with Outliers in High Dimension

### 2015-04-14 paper.tex
introduction

>>>>

The problem without outliers (i.e., $z=0$) in general (Euclidean) space can not be approximated with a smaller than factor $2$ ($1.822$) unless $\mathbf{P}=\mathbf{NP}$ \cite{kim:6}.
Gonzalez showed a $2$-approximation algorithm for this problem \cite{kim:17}.
For small $k$ and $d$, there are more efficient algorithms.
For Euclidean $1$-center in fixed dimensions, Chazelle \etal{} gave a linear algorithm \cite{kim:12}.
For $2$-center in the plane the best known algorithm, runs in $\cO(n \log^2 n \log^2 \log n)$ and linear space, given by Chan \cite{kim:8}.

In streaming data model, Guha \cite{kim:18}, in parallel with McCutchen \etal{} \cite{kim:22} gave $(2+\eps)$-approximation for $k$-center without outliers.
For small $k$, better algorithms are known. For fixed $d$, there exists an $\eps$-coreset for $1$-center with $\cO(1 / \eps^{\frac{d-1}{2}})$ space.
For higher dimensions, Agarwal gave a $\frac{1+\sqrt{3}}{2}$-approximation for $1$-center problem, and Chan \cite{kim:10} reduced its approximation factor to $1.22$, by an improved analysis.
For $k=2$, Ahn \etal{} \cite{ahn14} gave a $(2+\eps)$-approximation algorithm using $\cO(\frac{d}{\eps})$ space, which was improved for Euclidean spaces by Kim \etal{} \cite{kim14} to a $(1.8+\eps)$-approximation algorithm, using the same space.

For the problem of $k$-center with outliers, Charikar \etal{} \cite{zarrabi:8} gives the first offline algorithm with approximation factor $3$ ($4$), where the candidate center points of the centers can be enumerated (otherwise).
For small $k$, there are better results known. $1$-center with $z$ outliers can be solved in $\cO(n \log z + z^{3} n^{\eps})$, for any $\eps > 0$, using Matou{\v{s}}ek's framework \cite{on-geometric-optimization-with-few-violated-constraints}.
Agarwal \cite{an efficient algorithm for 2d euclidean 2-center with outliers} gives an algorithm with running time $\cO(n z^{7} \log^3 z)$ for $2$-center problem in $\IR^2$.

For the problem of $k$-center with $z$ outliers in streaming data model, McCutchen \etal{} \cite{zarrabi:14} (using idea in Charikar \etal{} \cite{zarrabi:8} as subroutine) give a $(3+\eps)$-approximation ($(4+\eps)$-approximation, depending on the used subroutine) algorithm using $\cO(\frac{zk}{\eps})$ space.
In special case of $k=1$ in fixed dimensions, there exists a streaming algorithm with $\cO(\frac{z}{\eps^{\cO(d)}})$ space \cite{zarrabi:4, zarrabi:15}.
For Euclidean spaces, Zarrabi-Zadeh gave $(1.22 \sqrt{2})$\nobreakdash-approximation for $1$-center with $z$ outliers running in $\cO(?)$ space.

In this paper we first give a simple $2$-approximation algorithm for $1$-center problem with space complexity $\cO(z^2 + zd)$. We then show a $(1.8 + \eps)$-approximation streaming algorithm for $2$-center with $z$ outliers in Euclidean space, with space complexity polynomial in $z$ and $d$ and sub-linear in $n$.
